---
title: "Homework 5"
author: "Yuki Joyama"
date: "2023-11-07"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1 
```{r message=FALSE}
library(tidyverse)

# dataset preparation
df_homicide = read_csv("./data/homicide-data.csv") |> 
  janitor::clean_names() 
```

The homicide data collected by the Washington Post has `r ncol(df_homicide)` columns and `r nrow(df_homicide)` rows. The data set comprises information of homicide such as incident dates, victim names, their race, gender, location, etc.

```{r message=FALSE}
df_homicide = df_homicide |>  
  mutate(
    city_state = paste(city, state, sep = ", ") # create a city_state variable
  ) 

# the total number of homicides within cities
df_n_homicide = df_homicide |> 
  group_by(city_state) |> 
  summarise(n_homicide = n()) |> 
  arrange(desc(n_homicide))

# the number of unsolved homicides within cities
df_n_unsolved_homicide = df_homicide |> 
  filter(disposition == "Closed without arrest" | disposition == "Open/No arrest") |> 
  group_by(city_state) |> 
  summarise(n_unsolved_homicide = n()) |> 
  arrange(desc(n_unsolved_homicide))

# combine two data sets
df_n_homicide = left_join(df_n_homicide, df_n_unsolved_homicide) |> 
  drop_na()

df_n_homicide
```

The above outputs show the total number of homicides and unsolved homicides within cities. We can see that Chicago, IL has the most homicide cases. 

```{r}
# estimate the proportion of homicides that are unsolved in Baltimore, MD
prop.test(x = filter(df_n_homicide, city_state == "Baltimore, MD") |> pull(n_unsolved_homicide), n = filter(df_n_homicide, city_state == "Baltimore, MD") |> pull(n_homicide)) |> 
  broom::tidy() |> 
  knitr::kable(digits = 2)
```

The estimated proportion of homicides that are unsolved in Baltimore, MD is 0.65 with 95% confidence interval 0.63 - 0.66.

```{r}
# run prop.test for each of the cities
unsolved_prop_test = df_n_homicide |> 
  mutate(
    prop_test = purrr::map2(pull(df_n_homicide, n_unsolved_homicide), pull(df_n_homicide, n_homicide), \(x, y) prop.test(x = x, n = y)),
    tidy_test = purrr::map(prop_test, broom::tidy) 
  ) |> 
  unnest(tidy_test) |> 
  select(city_state, estimate, conf.low, conf.high)

# create a plot that shows the estimates and CIs for each city
unsolved_prop_test |> 
  mutate(city_state = fct_reorder(city_state, estimate)) |> 
  ggplot(aes(x = city_state, y = estimate)) +
  geom_point() +
  geom_errorbar(aes(x = city_state, ymin = conf.low, ymax = conf.high)) +
  theme(axis.text.x = element_text(angle = 90))
```

The city with the lowest estimate of the proportion for unsolved homicide is Richmond, VA. The city with the highest estimate is Chicago, IL.


